Extended Abstract

### 2021-05-24 paper/paper.tex
Rework abstract and intro

\begin{abstract}
  While a \picalc{} with linear types ensures the privacy and safety of concurrent communication, allowing unrestricted non-linear communication is key to model real-world services.
  To implement a decision procedure that type checks a \picalc{} process with linear and shared types without relying on user-provided type annotations we must infer the types of the channels created within the process.
  If we limit ourselves to the shared \picalc{}, we can traverse the process bottom-up and eagerly solve typing constraints into substitutions and apply them to the typing context.
  However, in a setting with both linear and shared types, typing constraints do not always come with a most general solution, and thus cannot always be eagerly solved.

  We follow Padovani \cite{Padovani15} and provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up and, in addition to the typing context, collects a set of typing constraints.
  We then solve those constraints that have a most general solution (by using well-known unification algorithms \cite{McBride03}) while deferring the rest until more information becomes available.
  We state clear soundness and completeness theorems separating both these phases, and a progress theorem that ensures that only those constraints without a most general solution are deferred.
  The totality of this work is being mechanised in Agda.
\end{abstract}

@@

To type check a \picalc{} process with shared and linear types we must assign a type to every communication channel created within the process.
To do so we can either ask the user for type annotations \cite{ZalakainD21}, or we can infer types by looking at how channels are used.
We follow this latter approach: we traverse processes bottom-up while keeping a typing context with \emph{metavariables} (\emph{holes}) and collecting typing constraints on metavariables, in line with co-contextual typing \cite{ErdwegBKKM15}.
Keeping a strictly bottom-up information flow has the additional advantage of making typing inference easily parallelisable.
Constraints with a most general solution are solved into substitutions and applied to the typing context, constraints without one must be kept around for later: applying them as substitutions risks over-constraining the problem down the line.
Armed with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, type checking that $\Delta \vdash P$ for some $\Delta$ amounts to emitting the extra constraint $\Gamma = \Delta$, assuming the metavariables in $\Gamma$ and $\Delta$ are disjoint.

>>>>

\begin{abstract}
  A \picalc{} with linear types ensures the privacy and safety of concurrent communication.
  Allowing unrestricted non-linear communication however is key to model real-world services.
  To implement a decision procedure that type checks a \picalc{} process with both linear and shared types one can either rely on user-provided type annotations, or infer the types of the channels created within the process.
  We choose to reduce the burden on the user by following the latter approach.
  If we limit ourselves to the shared \picalc{}, we can traverse a process bottom-up and eagerly solve typing constraints into substitutions and apply them to the typing context.
  However, in a setting with both linear and shared types, typing constraints do not always come with a most general solution, and thus cannot always be eagerly solved.

  We provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up and, in addition to the typing context, collects a set of typing constraints.
  We then solve those constraints that have a most general solution (by using well-known unification algorithms \cite{McBride03}) while deferring the rest until more information becomes available.
  We state clear soundness and completeness theorems separating both these phases, and a progress theorem that ensures that only those constraints without a most general solution are deferred.
  The totality of this work is being mechanised in Agda.
\end{abstract}

@@

To type check a \picalc{} process with shared and linear types we must assign a type to every communication channel created within the process.
To do so we can either ask the user for type annotations \cite{ZalakainD21}, or we can infer types by looking at how channels are used.
To reduce the burden on the user, we follow the latter approach: we traverse processes bottom-up while keeping a typing context with \emph{metavariables} (\emph{holes}) and collecting typing constraints on metavariables, in line with co-contextual typing \cite{ErdwegBKKM15}.
Keeping a strictly bottom-up information flow has the additional advantage of making typing inference easily parallelisable.
Constraints with a most general solution are solved into substitutions and applied to the typing context, constraints without one must be kept around for later: applying them as substitutions risks over-constraining the problem down the line.
Armed with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, type checking that $\Delta \vdash P$ for some $\Delta$ amounts to emitting the extra constraint $\Gamma = \Delta$, assuming the metavariables in $\Gamma$ and $\Delta$ are disjoint.
